---
title: 'Lab 9: T test and ANOVA'
author: "Your name and student ID"
date: "today's date"
output: pdf_document
---


### Instructions 
* Due date: Tuesday, August 9 at 10:00pm PST with 2 hour grace period.
* Late penalty: 50% late penalty if submitted within 24 hours of due date, no 
marks for assignments submitted thereafter.
* This assignment is graded on **correct completion**, all or nothing. You must pass all public tests and submit the assignment for credit.
* Submission process: Follow the submission instructions on the final page. Make sure you do not remove any `\newpage` tags or rename this file, as this will break the submission.


### Part 1: T test and NHANES
The NHANES is a large national survey conducted by the CDC.  Here we will look at a reduced set of data from the NHANES

```{r read, echo=FALSE}
# load in the appropriate libraries
library(readr)
library(ggplot2)
library(dplyr)
library(broom)
library(testthat)

# Read CSV into R
# include data/ because the file is in the data folder
nhanes <- read_csv(file="data/nhanes.csv") 

nhanes <- na.omit(nhanes) #this code skip rows with missing values
```

1. [ 1 mark] We are interested in looking at the systolic blood pressure "bpxsy1" by history of smoking "hs"
Start by generating an appropriate box plot to look at these data.

```{r box, echo=TRUE}
plot1 <- NULL # YOUR CODE HERE
plot1
```

```{r}
. = ottr::check("tests/p1.R")
```

\newpage

2. [1 mark] Now generate a set of faceted histograms that show the same data.

```{r hist, echo=TRUE}
plot2 <- 'Your plot here'
plot2

```

```{r}
. = ottr::check("tests/p2.R")
```


\newpage

3. [2 marks] Summarize the means and standard deviations of the systolic blood pressure for each category of `hs` using dplyr functions. `p3` should end up being a dataframe - check this in your environment.

```{r hist2, echo=TRUE}
p3 <- NULL # YOUR CODE HERE
p3

```

```{r}
. = ottr::check("tests/p3.R")
```

\newpage

4. [1 mark] Now consider the assumptions that need to be hold in order to run the two-sample t-test. Do they hold here? Why or why not?

_Type your answer here, replacing this text._


\newpage

5. [2 marks] State your null and alternative hypotheses.

_Type your answer here, replacing this text._


\newpage

6. [2 marks] Now run the t-test to see if the variability gives us evidence to reject the null hypothesis of no difference between blood pressure means by smoking history.

```{r,echo=TRUE}
p6 <- NULL # YOUR CODE HERE
p6
```

```{r}
. = ottr::check("tests/p6.R")
```

\newpage

7.  [2 marks] Use these results to interpret your p-value in context. Following that, decide whether to accept or reject the null hypothesis.

_Type your answer here, replacing this text._


\newpage

8. [2 marks] Repeat this analysis without using the t.test function.

First, you will need to get your test statistic:

```{r}
#this code gives you the number of smokers in the dataset
n_s <- nrow(nhanes %>% filter(hs == 'History of smoking'))
n_s

#this code gives you the number of non-smokers in the dataset
n_ns <- nrow(nhanes %>% filter(hs == 'No'))
n_ns

#calculate your test statistic. You can make more objects if you wish.
t_stat <- 'Your code here'
t_stat

n_ns <- nrow(nhanes %>% filter(hs == 'No'))

```

```{r}
. = ottr::check("tests/p8-1.R")
```

Now compare your test statistic to a t-distribution with df = 558 and calculate the p-value. This is an approximation using the smaller of the two sample sizes - 1.

```{r,echo=TRUE}
df <- n_ns - 1
df

p_value <- 'Your code here'
p_value

df <- n_ns - 1

```

```{r}
. = ottr::check("tests/p8-2.R")
```


\newpage

9. [2 marks] Finally, construct a 99% confidence interval for these data. Interpret the interval in context and decide whether or not to reject the null hypothesis based on these data.

```{r}
lowerbound <- 'Input your lower bound here'
upperbound <- 'Input your upper bound here'
conf_int <- c(lowerbound, upperbound)
conf_int

```

```{r}
. = ottr::check("tests/p9.R")
```
_Type your answer here, replacing this text._

\newpage

## Part 2: ANOVA

10. [ 1 mark] We are interested in looking at the systolic blood pressure "bpxsy1" by BMI category "bmicat"
Start by generating an appropriate box plot to look at these data.

```{r box-aov, echo=TRUE}
plot10 <- NULL # YOUR CODE HERE
plot10
```

```{r}
. = ottr::check("tests/p10.R")
```


\newpage

11. [1 mark] Now generate a set of faceted histograms that show the same data. It might be useful to assign a fill color to each category.

```{r hist-aov, echo=TRUE}
plot11 <- "Your plot here"
plot11

```

```{r}
. = ottr::check("tests/p11.R")
```
\newpage

12. [2 marks] Summarize the means and standard deviations of the outcome for each BMI category

```{r hist2-aov, echo=TRUE}
p12 <- "Your code here"
p12

```

```{r}
. = ottr::check("tests/p12.R")
```

\newpage

13. [2 marks] Now run the ANOVA test to see if the variability gives us evidence to reject the null hypothesis of no difference between blood pressure means by BMI category.

```{r,echo=TRUE}
p13 <- 'YOUR CODE HERE'
tidy(p13) #tidy displays your output. It lives in the `broom` package

```

```{r}
. = ottr::check("tests/p13.R")
```

\newpage

14.  [2 marks] Use these results to conduct a Tukey's HSD for these groups. Use the standard error rate of 5%. What conclusion can you draw?

```{r,echo=TRUE}
p14 <- NULL # YOUR CODE HERE
tidy(p14)
```

```{r}
. = ottr::check("tests/p14.R")
```
_Type your answer here, replacing this text._

\newpage

### Submission

For assignments in this class, you'll be submitting using the **Terminal** tab in the pane below. In order for the submission to work properly, make sure that:

1. Any image files you add that are needed to knit the file are in the `src` folder and file paths are specified accordingly. 
2. You **have not changed the file name** of the assignment.
3. The file is saved (the file name in the tab should be **black**, not red with an asterisk).
4. The file knits properly.

Once you have checked these items, you can proceed to submit your assignment.

1. Click on the **Terminal** tab in the pane below.
2. Copy-paste the following line of code into the terminal and press enter.

cd; cd ph142-su21/lab/lab09; python3 turn_in.py

3. Follow the prompts to enter your Gradescope username and password. When entering your password, you won't see anything come up on the screen--don't worry! This is just for security purposes--just keep typing and hit enter.
4. If the submission is successful, you should see "Submission successful!" appear as output.
5. If the submission fails, try to diagnose the issue using the error messages--if you have problems, post on Piazza. 

The late policy will be strictly enforced, **no matter the reason**, including submission issues, so be sure to submit early enough to have time to diagnose issues if problems arise.

