---
title: "Homework 11"
author: "Your name and student ID"
date: "Today's date"
output:
  pdf_document: default
---


### Instructions 
* Solutions will be released on August 10
* This semester, homework assignments are for practice only and will not be turned in for marks.

```{r, warning=F, message=F, echo=F}
library(dplyr)
library(ggplot2)
library(tidyverse)
```

Helpful hints:

-  Every function you need to use was taught during lecture! So you may need to 
revisit the lecture code to help you along by opening the relevant files on Datahub. 
Alternatively, you may wish to view the code in the condensed PDFs posted 
on the course website. Good luck!

- Knit your file early and often to minimize knitting errors! If you copy and 
paste code for the slides, you are bound to get an error that is hard to 
diagnose. Typing out the code is the way to smooth knitting! We recommend 
knitting your file each time after you write a few sentences/add a new code 
chunk, so you can detect the source of knitting errors more easily. This will 
save you and the GSIs from frustration! **You must knit correctly before submitting.**

- If your code runs off the page of the knitted PDF then you will LOSE POINTS! To
avoid this, have a look at your knitted PDF and ensure all the code fits in the 
file (you can easily view it on Gradescope via the provided link after submitting). 
If it doesn't look right, go back to your .Rmd file and add spaces (new lines) using 
the return or enter key so that the code runs onto the next line.

\newpage


Parental leave is often compensated to some degree, but the amount
of compensation varies greatly. You read a research article that stated, 
"across people of all incomes, 47% of leave-takers received full pay during 
their leave, 16% received partial pay, and 37% received no pay." 

After reading this, you wonder what the distribution of parental leave payment is for low income households. Suppose you conduct a survey of leave-takers within households earning less than $30,000 per year. You surveyed 225 people (selected in a random sample) and found that 51 received full pay, 33 received partial pay, and 141 received no pay.

1. [1 point] You would like to investigate whether the distribution of pay for households earning < $30,000 is different from that of all income levels. Does this correspond to a chi-square test of independence or a chi-square test for goodness of fit?

_Type your answer here, replacing this text._

\newpage


2. [1 point] What are the expected counts of leave-takers among households with incomes < $30,000? Assign each expected count to the appropriate variable. Make sure to remove the quotes. Round each number to 2 decimal places.

```{r}
# put your answer here
full_pay <- "REPLACE WITH NUMBER ROUNDED TO 2 DECIMALS"
partial_pay <- "REPLACE WITH NUMBER ROUNDED TO 2 DECIMALS"
no_pay <- "REPLACE WITH NUMBER ROUNDED TO 2 DECIMALS"

```

```{r}
. = ottr::check("tests/p2.R")
```


\newpage

3. [1 point] State the null hypothesis under which the above expected counts were computed.

_Type your answer here, replacing this text._

\newpage


4. [1 point] Compute the chi-square statistic. Round your answer to 2 decimal places.

```{r}
# put your answer here
chi_sq_answer <- "REPLACE WITH NUMBER ROUNDED TO 2 DECIMALS"

```

```{r}
. = ottr::check("tests/p4.R")
```


\newpage

5. [1 point] Uncomment which cell (i.e. term in summation) contributes the most to the statistic. 

```{r}
# UNCOMMENT THE CORRECT ANSWER

# largest_contribution <- "full pay"
# largest_contribution <- "partial pay"
# largest_contribution <- "no pay"

```

```{r}
. = ottr::check("tests/p5.R")
```

\newpage

6. [1 point] Compute the p-value for your test statistic. Round your answer to 2 decimal places.

```{r}
# put your answer here
p_value <- "REPLACE WITH NUMBER ROUNDED TO 2 DECIMALS"

```

```{r}
. = ottr::check("tests/p6.R")
```
\newpage

7. [1 point] Conclude whether you believe there is evidence against the null hypothesis in favor of the alternative hypothesis. Answer this by uncommenting the appropriate conclusion.

```{r}
# UNCOMMENT THE CORRECT ANSWER

# conclusion <- "in favor of null"
# conclusion <- "against null"

```

```{r}
. = ottr::check("tests/p7.R")
```


\newpage

Human papillomavirus (HPV) is a very common STI that most sexually
active persons will encounter during their lifetimes. While many people clear the virus, certain strands can lead to adverse health outcomes such as genital warts and cervical cancer. 

Suppose that you selected a random sample from a population and collected
these data on age and HPV status for the sample:

|Age Group    | HPV +        | HPV -          | Row total   |
|:-----------:|:------------:|:--------------:|:-----------:|
|14-19        | 160          | 492            | 652 (33.9%) |
|20-24        | 85           | 104            | 189 (9.8%)  |
|25-29        | 48           | 126            | 174 (9.1%)  |
|30-39        | 90           | 238            | 328 (17.1%) |
|40-49        | 82           | 242            | 324 (16.9%) |
|50-59        | 50           | 204            | 254 (13.2%) |
| Col total   | 515 (26.8%)  | 1406 (73.2%)   | 1921        |

8. [1 point] Which variable is explanatory and which is response? Uncomment the appropriate answer.

```{r}
# UNCOMMENT THE CORRECT ANSWER

# variable_type <- c("explanatory: age group", "response: HPV status")
# variable_type <- c("explanatory: HPV status", "response: age group")

```

```{r}
. = ottr::check("tests/p8.R")
```

\newpage

9. [2 points] Formulate null and alternative hypotheses using these data to test whether there is a relationship between age group and HPV status. State these hypotheses using the language or notation of conditional distributions.

_Type your answer here, replacing this text._


\newpage

10. [1 point] Run the code below to examine the conditional distribution of age by HPV status. Based on this plot, which age group will contribute the most to the chi-square statistic? Explain why.
(That is, can you tell based on this plot when the observed count will differ most from the expected count under the null hypothesis of no relationship between age group and HPV status?) 

```{r create-hpv.data, echo=F, out.width="80%", fig.align=T, message=F}
age_group <- rep(c("14-19", "20-24", "25-29", "30-39", "40-49", "50-59"), 2)
hpv_test <- c(rep("positive", 6), rep("negative", 6))
number_of_women <- c(160, 85, 48, 90, 82, 50, 
                     492, 104, 126, 238, 242, 204)

hpv_data <- data.frame(age_group, hpv_test, number_of_women)

hpv_data <- hpv_data %>% group_by(hpv_test) %>% 
  mutate(sum_women = sum(number_of_women),
         prop_women_age = number_of_women/sum_women)

ggplot(hpv_data, aes(x = age_group, y = prop_women_age)) + 
  geom_bar(aes(fill = hpv_test), stat = "identity", position = "dodge") +
  theme_minimal(base_size = 15) +
  labs(y = "Proportion", x = "Age", 
       main = "Conditional distribution of age by HPV status")
```

_Type your answer here, replacing this text._


\newpage

11. [2 points] Fill out the table of expected counts under the null hypothesis of no association between age group and HPV status. You don't need to show your work, but make sure you can calculate the expected counts by hand, using a calculator. Assign each appropriate cell/letter to the variable in the code. Round each number to 2 decimal places.

Expected counts:

|Age Group    | HPV +                     | HPV -                     | 
|:-----------:|:-------------------------:|:-------------------------:|
|14-19        |          A                |          H                | 
|20-24        |          B                |          I                | 
|25-29        |          C                |          J                | 
|30-39        |          D                |          K                | 
|40-49        |          E                |          L                |
|50-59        |          G                |          M                |

```{r}
# put your answer here
A <- "REPLACE WITH NUMBER ROUNDED TO 2 DECIMALS"
B <- "REPLACE WITH NUMBER ROUNDED TO 2 DECIMALS"
C <- "REPLACE WITH NUMBER ROUNDED TO 2 DECIMALS"
D <- "REPLACE WITH NUMBER ROUNDED TO 2 DECIMALS"
E <- "REPLACE WITH NUMBER ROUNDED TO 2 DECIMALS"
G <- "REPLACE WITH NUMBER ROUNDED TO 2 DECIMALS"
H <- "REPLACE WITH NUMBER ROUNDED TO 2 DECIMALS"
I <- "REPLACE WITH NUMBER ROUNDED TO 2 DECIMALS"
J <- "REPLACE WITH NUMBER ROUNDED TO 2 DECIMALS"
K <- "REPLACE WITH NUMBER ROUNDED TO 2 DECIMALS"
L <- "REPLACE WITH NUMBER ROUNDED TO 2 DECIMALS"
M <- "REPLACE WITH NUMBER ROUNDED TO 2 DECIMALS"

```

```{r}
. = ottr::check("tests/p11.R")
```


\newpage

12. [1 point] Calculate the test statistic. Round your answer to 2 decimal places.

```{r}
# put your answer here
chi_sq_p12 <- "REPLACE WITH NUMBER ROUNDED TO 2 DECIMALS"

```

```{r}
. = ottr::check("tests/p12.R")
```


\newpage

13. [1 point] Calculate the p-value for your test statistic. Round your answer to 2 decimal places.

```{r}
p_value_p13 <- "REPLACE WITH NUMBER ROUNDED TO 2 DECIMALS"

```

```{r}
. = ottr::check("tests/p13.R")
```

\newpage

14. [1 point] Assess whether there is evidence against the null in favor of the alternative. Answer this by uncommenting the appropriate conclusion.

```{r}
# UNCOMMENT THE CORRECT ANSWER
# conclusion_p14 <- "in favor of null"
# conclusion_p14 <- "against null"

```


