---
title: "Normal Distrubution"
author: "Your name and Student ID"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: pdf_document
---


### Instructions 
* Due date: Tuesday, July 27 at 10:00pm PST with 2 hour grace period.
* Late penalty: 50% late penalty if submitted within 24 hours of due date, no 
marks for assignments submitted thereafter.
* This assignment is graded on **correct completion**, all or nothing. You must pass all public tests and submit the assignment for credit.
* Submission process: Follow the submission instructions on the final page. Make sure you do not remove any `\newpage` tags or rename this file, as this will break the submission.

Helpful hints:

- Every function you need to use was taught during lecture! So you may need to 
revisit the lecture code to help you along by opening the relevant files on Datahub. 
Alternatively, you may wish to view the code in the condensed PDFs posted 
on the course website. Good luck!

- Knit your file early and often to minimize knitting errors! If you copy and 
paste code for the slides, you are bound to get an error that is hard to 
diagnose. Typing out the code is the way to smooth knitting! We recommend 
knitting your file each time after you write a few sentences/add a new code 
chunk, so you can detect the source of knitting errors more easily. This will 
save you and the GSIs from frustration! **You must knit correctly before submitting.**

- If your code runs off the page of the knitted PDF then you will LOSE POINTS! To
avoid this, have a look at your knitted PDF and ensure all the code fits in the 
file (you can easily view it on Gradescope via the provided link after submitting). 
If it doesn't look right, go back to your .Rmd file and add spaces (new lines) using 
the return or enter key so that the code runs onto the next line.

\newpage

### Normal Distribution

We will be introducing the normal distribution in this lab.

Notation reminder: Baldi and Moore use a compact notation for specifying that a population has a distribution that follows a normal curve with mean $\mu$ and standard deviation $\sigma$ : N( $\mu$ , $\sigma$ )

R Functions Reminder: `qnorm` takes a probability as its input and gives back a value on the distribution (aka a z-score if the distribution is N(0,1)). By default, it assumes the probability area you enter (as a decimal, not a percent) is the area below (or less than) the z score you need.  The `pnorm` function takes a z value or an X value as its input and gives back a probability area. 

**Z scores. **
  (From Baldi and Moore, 3E question 11.27, 4E question 11.29) 
Use R to find the standardized value z that satisfies each of the following conditions. In each case, sketch a standard Normal curve with your value of z marked on the axis. You do not have to attach your diagrams. 
\newpage
**1. [1 point]	The probability is 0.8 that a randomly selected observation falls below z.**  

```{r Section II Q1a}
p1 <- NULL # YOUR CODE HERE
p1
```

```{r}
. = ottr::check("tests/p1.R")
```

\newpage 
**15. [1 point]	The probability is 0.35 that a randomly selected observation falls above z.** 

```{r Section II Q1b}
p2 <- NULL # YOUR CODE HERE
p2
```

```{r}
. = ottr::check("tests/p2.R")
```

**Full-Term Birth Weights (From Baldi and Moore, 3E question 11.31, 4E question 11.33) **

For babies born at full term (37 to 39 completed weeks of gestation), the distribution of birth weight (in grams) is approximately normally distributed with a mean of 3350 grams and a standard deviation of 440 grams, N(3350,440).


**3. [1 point] What is the 25th percentile of the birthweights for full term babies?** 

```{r Section II Q2a}
p3 <- NULL # YOUR CODE HERE
p3
```

```{r}
. = ottr::check("tests/p3.R")
```

**4. [1 point] What is the 90th percentile of the birthweights for full term babies?** 


```{r Section II Q2b}
p4 <- NULL # YOUR CODE HERE
p4
```

```{r}
. = ottr::check("tests/p4.R")
```


**5. [1 point] What is the range of the middle 50% of birthweights for full term babies? Write this as a vector containing two values to indicate the lowerbound and upperbound of the middle 50%.**


```{r Section II Q2c}
p5 <- NULL # YOUR CODE HERE
p5
```

```{r}
. = ottr::check("tests/p5.R")
```

**Newborn Respirations. (From Baldi and Moore, 3E questions 11.15-11.17, 4E question 11.17-11.19)**

**6. [1 point]** The respiratory rate per minute in newborns varies according to a distribution that is approximately Normal with mean 50 and standard deviation 5.
The probability (convert to a percentage and round to two decimal places) that a randomly chosen newborn has a respiratory rate of 55 per minute or more is approximately:

```{r Section II Q5a}
p6 <- NULL # YOUR CODE HERE
p6
```

```{r}
. = ottr::check("tests/p6.R")
```

**7. [1 point]** The probability (convert to a percentage and round to two decimal places) that a randomly chosen newborn has a respiratory rate per minute between 40 and 55 is approximately: 

```{r Section II Q5b}
p7 <- NULL # YOUR CODE HERE
p7
```

```{r}
. = ottr::check("tests/p7.R")
```

**Drosophila (From Baldi and Moore, 3E questions 11.20 and 11.22, 4E question 11.23)**

The common fruit fly, Drosophila melanogaster, is the most studied organism in genetic research because it is small, easy to grow, and reproduces rapidly. The length of the thorax (where the wings and legs attach) in a population of male fruit flies is approximately Normal with mean 0.800 millimeters (mm) and standard deviation 0.078 mm.

**8. [1 point]**	Choose a male fruit fly at random. The probability (convert to a percentage and round to two decimal places) that the fly you choose has a thorax longer than 1 mm is about:

```{r Section II Q6a}
p8 <- NULL # YOUR CODE HERE
p8
```

```{r}
. = ottr::check("tests/p8.R")
```


**Late Pre-Term Birth Weights (From Baldi and Moore, 3E question 11.32, 4E question 11.34)** 

How much of a difference do a couple of weeks make for birthweight? Late preterm babies are born with 35 to 37 weeks of completed gestation. The distribution of birth weight (in grams) or late preterm babies is approximately normally distributed with a mean of 2750 grams and a standard deviation of 560 grams,  N(2750,560).


**10. [1 point] What is the 25th percentile of the birthweights for late-preterm babies?** 

```{r Section II Q3a}
p10 <- NULL # YOUR CODE HERE
p10
```

```{r}
. = ottr::check("tests/p10.R")
```


**11. [1 point] What is the 90th percentile of the birthweights for late-preterm babies?**

```{r Section II Q3b}
p11 <- NULL # YOUR CODE HERE
p11
```

```{r}
. = ottr::check("tests/p11.R")
```


**12. [1 point] What is the range of the middle 50% of birthweights for late-preterm babies? Write your answer as a vector containing the lower and upperbounds of the middle 50%.**

```{r Section II Q3c}
p12 <- NULL # YOUR CODE HERE
p12
```

```{r}
. = ottr::check("tests/p12.R")
```


**13. [1 point] Compare your answers to the parts of full term babies (problems 14-18) to late-preterm babies. What do you notice?**

_Type your answer here, replacing this text._


**Z scores. (From Baldi and Moore, 3E question 11.25, 4E question 11.27)**

Use R’s pnorm function to find the proportion of observations from a standard Normal distribution that fall in each of the following regions. In each case, sketch a standard Normal curve and shade the area representing the region.You do not have to attach your diagrams.

**14. [1 point]**	z $\le$ -2.25
```{r Section II Q7a}
p14 <- NULL # YOUR CODE HERE
p14
```

```{r}
. = ottr::check("tests/p14.R")
```


**15. [1 point]**	z $\ge$ -2.25
```{r Section II Q7b}
p15 <- NULL # YOUR CODE HERE
p15
```

```{r}
. = ottr::check("tests/p15.R")
```

**16. [1 point]**	z > 1.77
```{r Section II Q7c}
p16 <- NULL # YOUR CODE HERE
p16
```

```{r}
. = ottr::check("tests/p16.R")
```

**17. [1 point]** -2.25 < z < 1.77
```{r Section II Q7d}
p17 <- NULL # YOUR CODE HERE
p17
```

```{r}
. = ottr::check("tests/p17.R")
```

\newpage

### Submission

For assignments in this class, you'll be submitting using the **Terminal** tab in the pane below. In order for the submission to work properly, make sure that:

1. Any image files you add that are needed to knit the file are in the `src` folder and file paths are specified accordingly. 
2. You **have not changed the file name** of the assignment.
3. The file is saved (the file name in the tab should be **black**, not red with an asterisk).
4. The file knits properly.

Once you have checked these items, you can proceed to submit your assignment.

1. Click on the **Terminal** tab in the pane below.
2. Copy-paste the following line of code into the terminal and press enter.

cd; cd ph142-su21/lab/lab06; python3 turn_in.py

3. Follow the prompts to enter your Gradescope username and password. When entering your password, you won't see anything come up on the screen--don't worry! This is just for security purposes--just keep typing and hit enter.
4. If the submission is successful, you should see "Submission successful!" appear as output.
5. If the submission fails, try to diagnose the issue using the error messages--if you have problems, post on Piazza. 

The late policy will be strictly enforced, **no matter the reason**, including submission issues, so be sure to submit early enough to have time to diagnose issues if problems arise.

